---
title: "metaoverfit: Introduction and Quick Start"
author: "Mahmood Ahmad"
date: "`r format(Sys.Date())`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{metaoverfit: Introduction and Quick Start}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse   = TRUE,
  comment    = "#>",
  fig.align  = "center",
  fig.width  = 6,
  fig.height = 4,
  message    = FALSE,
  warning    = FALSE
)
set.seed(123)
```

## Overview

**metaoverfit** provides tools to detect and correct overfitting in
meta-regression, including:

- Apparent and adjusted $R^2_{het}$
- Cross-validated (optimism-corrected) $\tilde{R}^2_{het}$
- Bootstrap confidence intervals with boundary-effect diagnostics
- Sample size guidance and diagnostic plots

This vignette walks through a minimal example using the BCG vaccine dataset
from the **metadat** package.

## Load Packages

```{r load-packages}
library(metaoverfit)
library(metadat)
library(metafor)
```

## Example: BCG Vaccine Trials

We use log risk ratios from the BCG dataset and test whether absolute
latitude explains between-study heterogeneity.

```{r prepare-data}
data(dat.bcg)
dat <- escalc(
  measure = "RR",
  ai = tpos, bi = tneg, ci = cpos, di = cneg,
  data = dat.bcg
)
head(dat[, c("yi", "vi")])
```

### 1) Quick Overfitting Check

`check_overfitting()` runs LOOCV and a bootstrap, returning a concise report.

```{r overfitting-check}
res <- suppressWarnings(
  check_overfitting(
    yi   = yi,
    vi   = vi,
    mods = ~ ablat,
    data = dat,
    B    = 200          # keep runtime low for vignettes
  )
)
print(res)
```

**Interpretation** (typical for small $k$):

- Apparent $R^2_{het}$ can be large.
- LOOCV-corrected $\tilde{R}^2_{het}$ is often much smaller.
- Bootstrap CIs commonly include 0, indicating weak evidence of genuine
  explanation.

### 2) Cross-Validated $\tilde{R}^2_{het}$

Compute CV metrics directly (LOOCV refits $k$ times):

```{r cv-direct}
cv_out <- suppressWarnings(
  r2het_cv(
    dat$yi, dat$vi,
    mods = cbind(1, dat.bcg$ablat),
    cv_method = "loo",
    verbose   = FALSE
  )
)
cv_out$r2het_apparent
cv_out$r2het_corrected
cv_out$optimism
```

### 3) Bootstrap Confidence Intervals

For the vignette we use a modest number of replicates. In research, increase
`B` to 1000--5000.

```{r bootstrap}
boot_out <- suppressWarnings(
  r2het_boot(
    dat$yi, dat$vi,
    mods = cbind(1, dat.bcg$ablat),
    B = 200,
    conf_level = 0.95
  )
)
boot_out$ci
boot_out$boundary_rate
```

### 4) Plots

A bar plot comparing apparent vs. corrected $R^2_{het}$:

```{r plot-bar, fig.cap = "Apparent vs. corrected R2het."}
plot_overfitting(cv_out, type = "bar")
```

A histogram of bootstrap replicates (requires a `metaoverfit_check` object):

```{r plot-histogram, fig.cap = "Bootstrap distribution of R2het."}
plot_overfitting(res, type = "histogram")
```

### 5) Sample Size Guidance

Plan studies with acceptable optimism:

```{r sample-size}
sample_size_recommendation(p = 2, target_optimism = 0.10)
```

## Tips and Notes

- Small $k$ and many moderators increase the chance of $\tau^2$ estimates at
  the boundary (0); `metafor::rma()` may warn. In CV/bootstrap, this is
  expected for small datasets and is handled internally.
- For reproducible reports, set a seed and increase `B` outside of
  vignettes.
- Boundary hits above 10% suggest the dataset is too sparse for the model.

## Session Info

```{r session-info}
sessionInfo()
```

## References

1. Viechtbauer W (2010). Conducting meta-analyses in R with the metafor
   package. *J Stat Softw*, 36:1--48.
2. White T, Viechtbauer W. *metadat: Meta-analytic datasets for R*.
3. See `?r2het_cv`, `?r2het_boot`, `?check_overfitting` for API docs.
